---
title: "Supported Plot Types in MAIDR"
author: "JooYoung Seo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Supported Plot Types in MAIDR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

MAIDR supports a comprehensive range of plot types from both **ggplot2** and **Base R**. This vignette demonstrates each supported visualization with example code.

## Bar Charts

### Simple Bar Chart

```{r simple-bar}
library(maidr)
library(ggplot2)

# ggplot2
bar_data <- data.frame(
  Category = c("A", "B", "C", "D"),
  Value = c(30, 25, 35, 20)
)

p <- ggplot(bar_data, aes(x = Category, y = Value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Simple Bar Chart") +
  theme_minimal()

show(p)

# Base R equivalent
barplot(bar_data$Value, 
        names.arg = bar_data$Category,
        col = "steelblue",
        main = "Simple Bar Chart")
show()  # Note: No arguments for Base R plots
```

### Dodged Bar Chart

```{r dodged-bar}
library(maidr)
library(ggplot2)

# ggplot2
dodged_data <- data.frame(
  Category = rep(c("A", "B", "C"), each = 2),
  Type = rep(c("Type1", "Type2"), 3),
  Value = c(10, 15, 20, 25, 30, 35)
)

p <- ggplot(dodged_data, aes(x = Category, y = Value, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(title = "Dodged Bar Chart") +
  theme_minimal()

show(p)

# Base R equivalent
matrix_data <- matrix(c(10, 15, 20, 25, 30, 35), nrow = 2)
barplot(matrix_data, 
        beside = TRUE,  # Creates dodged effect
        names.arg = c("A", "B", "C"),
        col = c("steelblue", "coral"),
        legend = c("Type1", "Type2"),
        main = "Dodged Bar Chart")
show()
```

### Stacked Bar Chart

```{r stacked-bar}
library(maidr)
library(ggplot2)

# ggplot2
stacked_data <- data.frame(
  Category = rep(c("A", "B", "C"), each = 2),
  Type = rep(c("Type1", "Type2"), 3),
  Value = c(10, 15, 20, 25, 30, 35)
)

p <- ggplot(stacked_data, aes(x = Category, y = Value, fill = Type)) +
  geom_bar(stat = "identity", position = position_stack()) +
  labs(title = "Stacked Bar Chart") +
  theme_minimal()

show(p)

# Base R equivalent
matrix_data <- matrix(c(10, 15, 20, 25, 30, 35), nrow = 2)
barplot(matrix_data, 
        beside = FALSE,  # Creates stacked effect
        names.arg = c("A", "B", "C"),
        col = c("steelblue", "coral"),
        legend = c("Type1", "Type2"),
        main = "Stacked Bar Chart")
show()
```

## Histograms

```{r histogram}
library(maidr)
library(ggplot2)

# ggplot2
hist_data <- data.frame(values = rnorm(1000, mean = 0, sd = 1))

p <- ggplot(hist_data, aes(x = values)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  labs(title = "Histogram", x = "Values", y = "Frequency") +
  theme_minimal()

show(p)

# Base R equivalent
hist(rnorm(1000, mean = 0, sd = 1),
     breaks = 30,
     col = "skyblue",
     border = "black",
     main = "Histogram",
     xlab = "Values",
     ylab = "Frequency")
show()
```

## Scatter/Point Plots

```{r scatter}
library(maidr)
library(ggplot2)

# ggplot2
scatter_data <- data.frame(
  x = rnorm(50),
  y = rnorm(50),
  group = sample(c("A", "B", "C"), 50, replace = TRUE)
)

p <- ggplot(scatter_data, aes(x = x, y = y, color = group)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(title = "Scatter Plot") +
  theme_minimal()

show(p)

# Base R equivalent
x <- rnorm(50)
y <- rnorm(50)
plot(x, y,
     pch = 19,
     col = rainbow(3)[as.numeric(factor(sample(c("A", "B", "C"), 50, replace = TRUE)))],
     main = "Scatter Plot")
show()
```

## Line Plots

### Single Line

```{r single-line}
library(maidr)
library(ggplot2)

# ggplot2
line_data <- data.frame(
  x = 1:10,
  y = c(2, 4, 1, 5, 3, 7, 6, 8, 9, 4)
)

p <- ggplot(line_data, aes(x = x, y = y)) +
  geom_line(color = "steelblue", linewidth = 1.5) +
  labs(title = "Single Line Plot") +
  theme_minimal()

show(p)

# Base R equivalent
x <- 1:10
y <- c(2, 4, 1, 5, 3, 7, 6, 8, 9, 4)
plot(x, y, type = "l", col = "steelblue", lwd = 2, main = "Single Line Plot")
show()
```

### Multiple Lines

```{r multiline}
library(maidr)
library(ggplot2)

# ggplot2
multiline_data <- data.frame(
  x = rep(1:10, 3),
  y = c(
    c(2, 4, 1, 5, 3, 7, 6, 8, 9, 4),  # Series 1
    c(1, 3, 5, 2, 4, 6, 8, 7, 5, 3),  # Series 2
    c(3, 1, 4, 6, 5, 2, 4, 5, 7, 6)   # Series 3
  ),
  series = rep(c("A", "B", "C"), each = 10)
)

p <- ggplot(multiline_data, aes(x = x, y = y, color = series)) +
  geom_line(linewidth = 1) +
  labs(title = "Multiline Plot") +
  theme_minimal()

show(p)

# Base R equivalent
x <- 1:10
y1 <- c(2, 4, 1, 5, 3, 7, 6, 8, 9, 4)
y2 <- c(1, 3, 5, 2, 4, 6, 8, 7, 5, 3)
y3 <- c(3, 1, 4, 6, 5, 2, 4, 5, 7, 6)

plot(x, y1, type = "l", col = "red", lwd = 2, main = "Multiline Plot", ylim = c(0, 10))
lines(x, y2, col = "blue", lwd = 2)
lines(x, y3, col = "green", lwd = 2)
show()
```

## Box Plots

```{r boxplot}
library(maidr)
library(ggplot2)

# ggplot2 - Horizontal
p <- ggplot(iris, aes(x = Petal.Length, y = Species)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7) +
  labs(title = "Boxplot - Petal Length by Species") +
  theme_minimal()

show(p)

# ggplot2 - Vertical
p_vert <- ggplot(iris, aes(x = Species, y = Petal.Length)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7) +
  labs(title = "Boxplot - Vertical Orientation") +
  theme_minimal()

show(p_vert)

# Base R equivalent
boxplot(Petal.Length ~ Species, 
        data = iris,
        col = "lightblue",
        main = "Boxplot - Petal Length by Species",
        horizontal = TRUE)
show()
```

## Heatmaps

```{r heatmap}
library(maidr)
library(ggplot2)

# ggplot2
heatmap_data <- data.frame(
  x = rep(c("A", "B"), each = 2),
  y = rep(c("1", "2"), 2),
  z = c(1, 2, 3, 4)
)

p <- ggplot(heatmap_data, aes(x = x, y = y, fill = z)) +
  geom_tile() +
  geom_text(aes(label = z), color = "white", size = 5) +
  labs(title = "Heatmap with Labels") +
  theme_minimal()

show(p)

# Base R equivalent
matrix_data <- matrix(c(1, 2, 3, 4), nrow = 2)
image(matrix_data, 
      col = heat.colors(10),
      main = "Heatmap",
      axes = FALSE)
axis(1, at = seq(0, 1, length.out = 2), labels = c("A", "B"))
axis(2, at = seq(0, 1, length.out = 2), labels = c("1", "2"))
show()
```

## Density/Smooth Plots

```{r density}
library(maidr)
library(ggplot2)

# ggplot2
density_data <- data.frame(values = rnorm(1000, mean = 0, sd = 1))

p <- ggplot(density_data, aes(x = values)) +
  geom_density(fill = "lightblue", alpha = 0.5) +
  labs(title = "Density Plot") +
  theme_minimal()

show(p)

# Base R equivalent
values <- rnorm(1000, mean = 0, sd = 1)
plot(density(values), 
     col = "blue",
     lwd = 2,
     main = "Density Plot")
polygon(density(values), col = rgb(0.678, 0.847, 0.902, 0.5))
show()
```

## Faceted Plots

```{r faceted}
library(maidr)
library(ggplot2)

# Faceted bar plot
facet_data <- data.frame(
  x = rep(1:5, 4),
  y = runif(20, 1, 100),
  group = rep(c("Group 1", "Group 2", "Group 3", "Group 4"), each = 5)
)

p <- ggplot(facet_data, aes(x = x, y = y)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  facet_wrap(~ group, ncol = 2) +
  labs(title = "Faceted Bar Plot") +
  theme_minimal()

show(p)
```

## Multi-Panel Layouts (Patchwork)

```{r patchwork}
library(maidr)
library(ggplot2)
library(patchwork)

# Create individual plots
p1 <- ggplot(data.frame(x = 1:10, y = rnorm(10)), aes(x, y)) +
  geom_line(color = "steelblue") +
  labs(title = "Line Plot") +
  theme_minimal()

p2 <- ggplot(data.frame(x = c("A", "B", "C"), y = c(10, 20, 15)), aes(x, y)) +
  geom_bar(stat = "identity", fill = "coral") +
  labs(title = "Bar Plot") +
  theme_minimal()

# Combine with patchwork
combined <- p1 + p2
show(combined)
```

## When to Use Each Plot Type

| Plot Type | Best For | Example Use Case |
|-----------|----------|------------------|
| **Bar Chart** | Comparing categories | Sales by product |
| **Histogram** | Showing distributions | Test score frequencies |
| **Scatter Plot** | Relationships between variables | Height vs weight |
| **Line Plot** | Trends over time/order | Stock prices |
| **Box Plot** | Distribution comparison | Salary by department |
| **Heatmap** | Matrix relationships | Correlation matrices |
| **Density** | Smooth distributions | Probability density |
| **Faceted** | Comparing subgroups | Regional sales trends |

## Next Steps

- **[Getting Started](getting-started.html)** - Learn the basics
- **[Shiny Integration](shiny-integration.html)** - Use in interactive apps
- **Package documentation** - Run `help(package = "maidr")`
