---
title: "Base R Time-Series and Diagnostic Examples"
description: "Accessible Base R time-series and model diagnostics with maidr: acf() correlograms, spectrum() and cpgram() periodograms, monthplot() seasonal subseries, lag plots, termplot() and Q-Q plots."
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include = FALSE}
library(maidr)
library(ggplot2)
maidr_on()
set.seed(42)
```

Base R's time-series and model-diagnostic charts have no ggplot2
equivalent, so maidr reads each by mapping it onto a layer type it already
knows: a correlogram becomes a lollipop walked lag by lag, a periodogram or
seasonal subseries a line, a lag plot or Q-Q plot a point layer, and the
Left and Right arrow keys and sonification then work exactly as they do
for those types. Every reading on this page is an experimental plot type
(see "Experimental Plot Types" in the README): none has been through a
user study and each may change without a deprecation period. The
[examples hub](examples.html) lists every other plot family.

> **Note:** Everything below is Base R, so no ggplot2 counterpart is
> shown. Each section names the reading its chart gets, because the mapping
> is not always the one the chart's name suggests: a correlogram is read as
> a lollipop, a cumulative periodogram as a step.

## Correlogram

`acf()`, `pacf()` and `ccf()` draw a spike per lag, and maidr reads each as
a **lollipop**: one term per lag, carrying the correlation at that lag. The
axes come from the function itself, so the reading says `Lag` against
`ACF`, `Partial ACF` or `CCF` without being told.

`acf()` starts at lag 0, whose autocorrelation is always exactly 1, so the
first term of an ACF is a constant rather than a measurement. `pacf()` has
no lag 0 and starts at lag 1.

```{r}
acf(lh, lag.max = 5, main = "Luteinizing hormone: autocorrelation")
```

```{r}
pacf(lh, lag.max = 5, main = "Luteinizing hormone: partial autocorrelation")
```

`ccf()` correlates two series across negative and positive lags, so its
terms are signed and the sign says which series leads.

```{r}
ccf(mdeaths, fdeaths, lag.max = 4, main = "Male vs female deaths")
```

## Spectral Density

`spectrum()` estimates how a series' variance is distributed across
frequencies. maidr reads the periodogram as a **line** of `spectrum`
against `frequency`.

```{r}
spectrum(lh, main = "Luteinizing hormone: spectral density")
```

## Cumulative Periodogram

`cpgram()` draws the cumulative periodogram against its confidence band.
The cumulative curve is read as a **step**; the band is drawing rather than
data, and is not announced.

> **Note:** `cpgram()` labels only its x axis, so the reading carries
> `frequency` and no y label. It takes no `ylab` argument to supply one.

```{r}
cpgram(lh, main = "Luteinizing hormone: cumulative periodogram")
```

## Seasonal Subseries

`monthplot()` breaks a seasonal series into one segment per cycle position,
so every January sits together and the seasonal shape is read directly.
maidr reads it as a **line**.

> **Note:** `monthplot()` has no default for `xlab` -- the axis it writes
> carries the cycle labels, not a quantity -- so x is named only when the
> call names it. Left out, y falls back to the series name — `nottem` here.

```{r}
monthplot(
  nottem,
  xlab = "Month", ylab = "Temperature (F)",
  main = "Nottingham temperatures by month"
)
```

## Lag Plot

`lag.plot()` plots a series against itself shifted by *k*, which is how
serial dependence is read by eye. maidr reads it as a **point** layer whose
x axis is named for the shift — `lag 1` against the series.

```{r}
lag.plot(lh, lags = 1, main = "Luteinizing hormone against its own lag")
```

## Partial Effects

`termplot()` draws each term's fitted contribution against its own
predictor, holding the rest fixed. maidr reads each panel as a **line**,
labelled `partial for <term>` so the reading says it is a contribution
rather than a fitted value.

```{r}
termplot(lm(mpg ~ wt, data = mtcars), main = "Partial effect of weight")
```

## Q-Q Plot

`qqnorm()` plots sample quantiles against theoretical ones, and `qqline()`
adds the reference line through the quartiles. Both are read — the points as
a **point** layer and the reference as a **line** — so the line a sighted
reader compares against is navigable rather than decoration.

```{r}
z <- rnorm(50)
qqnorm(z, main = "Normal Q-Q plot")
qqline(z)
```
