---
title: "Examples"
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include = FALSE}
library(maidr)
library(ggplot2)
maidr_on()
set.seed(42)
```

Making accessible data representation with **maidr** is easy and
straightforward. If you already have data visualization code using
**ggplot2** or **Base R**, you can make your plots accessible with maidr
in just a few lines of code.

Simply load the maidr package and call `maidr_on()`. When plots are
printed in a document, maidr will automatically generate accessible
versions. You can then interact with the accessible versions using
keyboard shortcuts (refer to the table below).

| Key | Action |
|:---:|--------|
| **Left / Right** | Navigate between data points |
| **B** | Toggle braille mode |
| **T** | Toggle text mode |
| **S** | Toggle sonification (hear the data) |
| **R** | Toggle review mode |

## When to Use Each Plot Type

| Plot Type | Best For | Example Use Case |
|-----------|----------|------------------|
| **Bar Chart** | Comparing categories | Sales by product |
| **Histogram** | Showing distributions | Test score frequencies |
| **Scatter Plot** | Relationships between variables | Height vs weight |
| **Line Plot** | Trends over time/order | Stock prices |
| **Box Plot** | Distribution comparison | Salary by department |
| **Violin Plot** | Distribution shape comparison | Gene expression by group |
| **Candlestick (OHLC)** | Financial OHLC time-series | Stock price movements |
| **Heatmap** | Matrix relationships | Correlation matrices |
| **Density** | Smooth distributions | Probability density |
| **Faceted** | Comparing subgroups | Regional sales trends |
| **Multi-Panel** | Multiple related views | Dashboard layouts |
| **Multi-Layered** | Combining visualizations | Histogram + density overlay |

---

## Bar Plots

### Simple Bar Chart

A simple bar chart compares values across categories. Each bar
represents one category with its height proportional to its value.

#### ggplot2

```{r}
bar_data <- data.frame(
  Product = c("Laptop", "Tablet", "Phone", "Monitor"),
  Sales = c(150, 230, 180, 290)
)

p <- ggplot(bar_data, aes(x = Product, y = Sales)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Product Sales by Category",
    x = "Product",
    y = "Sales (units)"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
products <- c("Laptop", "Tablet", "Phone", "Monitor")
sales <- c(150, 230, 180, 290)

barplot(sales,
  names.arg = products,
  col = "steelblue",
  main = "Product Sales by Category",
  xlab = "Product",
  ylab = "Sales (units)"
)
```

### Dodged / Grouped Bar Chart

A dodged bar chart places bars for each sub-group side by side, making
it easy to compare values within and across categories.

#### ggplot2

```{r}
dodged_data <- data.frame(
  Region = rep(c("North", "South", "East"), each = 2),
  Quarter = rep(c("Q1", "Q2"), 3),
  Revenue = c(120, 150, 200, 180, 160, 210)
)

p <- ggplot(dodged_data, aes(x = Region, y = Revenue, fill = Quarter)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(title = "Quarterly Revenue by Region") +
  scale_fill_manual(values = c("steelblue", "coral")) +
  theme_minimal()

p
```

#### Base R

```{r}
revenue_matrix <- matrix(c(120, 150, 200, 180, 160, 210), nrow = 2)
rownames(revenue_matrix) <- c("Q1", "Q2")
barplot(revenue_matrix,
  beside = TRUE,
  names.arg = c("North", "South", "East"),
  col = c("steelblue", "coral"),
  legend.text = rownames(revenue_matrix),
  main = "Quarterly Revenue by Region",
  xlab = "Region",
  ylab = "Revenue"
)
```

### Stacked Bar Chart

A stacked bar chart layers sub-groups on top of each other within each
category, showing both individual contributions and the total.

#### ggplot2

```{r}
stacked_data <- data.frame(
  Year = rep(c("2022", "2023", "2024"), each = 3),
  Source = rep(c("Solar", "Wind", "Hydro"), 3),
  Output = c(40, 30, 50, 55, 45, 48, 70, 60, 52)
)

p <- ggplot(stacked_data, aes(x = Year, y = Output, fill = Source)) +
  geom_bar(stat = "identity", position = position_stack()) +
  labs(
    title = "Renewable Energy Output by Source",
    y = "Output (GWh)"
  ) +
  scale_fill_manual(values = c("#2ecc71", "#3498db", "#9b59b6")) +
  theme_minimal()

p
```

#### Base R

```{r}
energy_matrix <- matrix(
  c(40, 30, 50, 55, 45, 48, 70, 60, 52),
  nrow = 3
)
rownames(energy_matrix) <- c("Solar", "Wind", "Hydro")
barplot(energy_matrix,
  beside = FALSE,
  names.arg = c("2022", "2023", "2024"),
  col = c("#2ecc71", "#3498db", "#9b59b6"),
  legend.text = rownames(energy_matrix),
  main = "Renewable Energy Output by Source",
  xlab = "Year",
  ylab = "Output (GWh)"
)
```

---

## Histogram

A histogram shows the frequency distribution of a continuous variable by
grouping values into bins. The height of each bar indicates how many
observations fall within that range.

#### ggplot2

```{r}
exam_scores <- data.frame(score = c(
  rnorm(500, mean = 72, sd = 10),
  rnorm(500, mean = 85, sd = 8)
))

p <- ggplot(exam_scores, aes(x = score)) +
  geom_histogram(bins = 25, fill = "skyblue", color = "white") +
  labs(
    title = "Distribution of Exam Scores",
    x = "Score",
    y = "Frequency"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
scores <- c(rnorm(500, mean = 72, sd = 10), rnorm(500, mean = 85, sd = 8))
hist(scores,
  breaks = 25,
  col = "skyblue",
  border = "white",
  main = "Distribution of Exam Scores",
  xlab = "Score",
  ylab = "Frequency"
)
```

---

## Scatter Plot

A scatter plot displays the relationship between two continuous
variables. Each point represents one observation, and color can encode a
third categorical variable.

#### ggplot2

```{r}
car_data <- data.frame(
  weight = mtcars$wt,
  mpg = mtcars$mpg,
  cylinders = factor(mtcars$cyl)
)

p <- ggplot(car_data, aes(x = weight, y = mpg, color = cylinders)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    title = "Fuel Efficiency vs Vehicle Weight",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon",
    color = "Cylinders"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
colors <- c("4" = "steelblue", "6" = "coral", "8" = "forestgreen")
plot(mtcars$wt, mtcars$mpg,
  pch = 19,
  col = colors[as.character(mtcars$cyl)],
  main = "Fuel Efficiency vs Vehicle Weight",
  xlab = "Weight (1000 lbs)",
  ylab = "Miles per Gallon"
)
legend("topright",
  legend = c("4 cyl", "6 cyl", "8 cyl"),
  col = colors, pch = 19
)
```

---

## Line Plots

### Single Line

A line plot connects ordered data points, making it ideal for showing
trends over time or sequential categories.

#### ggplot2

```{r}
temp_data <- data.frame(
  Month = factor(month.abb, levels = month.abb),
  Temperature = c(2, 4, 10, 15, 20, 25, 28, 27, 22, 15, 8, 3)
)

p <- ggplot(temp_data, aes(x = Month, y = Temperature, group = 1)) +
  geom_line(color = "tomato", linewidth = 1.2) +
  labs(
    title = "Average Monthly Temperature",
    y = "Temperature (C)"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
months <- 1:12
temps <- c(2, 4, 10, 15, 20, 25, 28, 27, 22, 15, 8, 3)
plot(months, temps,
  type = "l", col = "tomato", lwd = 2,
  main = "Average Monthly Temperature",
  xlab = "Month", ylab = "Temperature (C)",
  xaxt = "n"
)
axis(1, at = 1:12, labels = month.abb)
```

### Multiple Lines

A multi-line plot overlays several series on the same axes, enabling
direct comparison of trends across groups.

#### ggplot2

```{r}
multi_line <- data.frame(
  Year = rep(2015:2024, 3),
  Users = c(
    10, 15, 22, 35, 50, 72, 95, 120, 150, 180,
    8, 12, 18, 25, 38, 55, 70, 88, 110, 135,
    5, 8, 14, 20, 30, 42, 58, 75, 95, 118
  ),
  Platform = rep(c("Mobile", "Desktop", "Tablet"), each = 10)
)

p <- ggplot(multi_line, aes(x = Year, y = Users, color = Platform)) +
  geom_line(linewidth = 1.2) +
  labs(
    title = "Platform Users Over Time",
    y = "Users (millions)"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
years <- 2015:2024
users <- cbind(
  Mobile  = c(10, 15, 22, 35, 50, 72, 95, 120, 150, 180),
  Desktop = c(8, 12, 18, 25, 38, 55, 70, 88, 110, 135),
  Tablet  = c(5, 8, 14, 20, 30, 42, 58, 75, 95, 118)
)

matplot(years, users,
  type = "l", lwd = 2,
  col = c("steelblue", "coral", "forestgreen"),
  lty = 1,
  main = "Platform Users Over Time",
  xlab = "Year", ylab = "Users (millions)"
)
legend("topleft",
  legend = colnames(users),
  col = c("steelblue", "coral", "forestgreen"),
  lwd = 2
)
```

---

## Box Plot

A box plot summarizes a distribution using its five-number summary:
minimum, first quartile (Q1), median (Q2), third quartile (Q3), and
maximum. Outliers appear as individual points.

#### ggplot2

```{r}
p <- ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7) +
  labs(
    title = "Sepal Length by Iris Species",
    x = "Species",
    y = "Sepal Length (cm)"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
boxplot(Sepal.Length ~ Species,
  data = iris,
  col = "lightblue",
  main = "Sepal Length by Iris Species",
  xlab = "Species",
  ylab = "Sepal Length (cm)"
)
```

---

## Violin Plot

A violin plot combines kernel density estimation (KDE) curves with
box-summary statistics, providing a richer view of the data distribution
than a box plot alone. MAIDR renders each violin as two navigable
layers: a **box** layer (min, Q1, median, Q3, max) and a **KDE** layer
(density curve).

> **Note:** Violin plots are a ggplot2-only feature.

```{r}
p <- ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_violin(fill = "lightblue", alpha = 0.7) +
  labs(
    title = "MPG Distribution by Cylinder Count",
    x = "Cylinders",
    y = "Miles per Gallon"
  ) +
  theme_minimal()

p
```

---

## Candlestick (OHLC) Charts

A candlestick chart visualizes Open-High-Low-Close (OHLC) financial data.
Each candle represents one trading period and exposes the four price
fields plus a computed **trend** (Bull / Bear / Neutral) and
**volatility** (high − low). When volume data is present and combined
with a volume panel via patchwork, each candle's `volume` is also
embedded in the data point.

> **Note:** Requires the {tidyquant} package for the ggplot2 path and
> {quantmod} for the Base R path.
> Install with `install.packages(c("tidyquant", "patchwork", "quantmod"))`.

### Support matrix: what works in each system

The accessible HTML pipeline supports different sets of overlays for
the ggplot2 and Base R candlestick paths. Use the ggplot2 +
tidyquant + patchwork path whenever you need moving averages or a
volume sub-panel.

| Feature | ggplot2 (`tidyquant::geom_candlestick`) | Base R (`quantmod::chartSeries`) |
|---|---|---|
| Plain OHLC candlestick | ✅ Supported | ✅ Supported (OHLC-only input) |
| Moving-average overlay | ✅ via `tidyquant::geom_ma()` (one or more layers; auto-collapsed into a single multi-series line layer) | ❌ `TA = "addSMA()"` / `"addEMA()"` not supported |
| Volume sub-panel | ✅ via separate `geom_col()` + `patchwork::plot_layout()` (collapsed into the candlestick subplot, with `volume` embedded into each candle point) | ❌ `TA = "addVo()"` not supported; default `TA` with a `Volume` column also unsupported |
| Behavior when unsupported | n/a | One-time warning + fall back to native (non-accessible) graphics; advisory points users to the ggplot2 pipeline |

### Simple OHLC Candlestick

#### ggplot2

```{r}
library(tidyquant)

ohlc_simple <- data.frame(
  date  = as.Date(c("2023-01-02", "2023-01-03", "2023-01-04", "2023-01-05")),
  open  = c(100, 105, 110, 108),
  high  = c(115, 108, 112, 110),
  low   = c(95, 102, 105, 100),
  # Bull, Bear, Bull, Neutral
  close = c(110, 103, 111, 108)
)

p <- ggplot(
  ohlc_simple,
  aes(x = date, open = open, high = high, low = low, close = close)
) +
  geom_candlestick(
    colour_up = "darkgreen", colour_down = "red",
    fill_up = "darkgreen", fill_down = "red"
  ) +
  labs(
    title = "Sample OHLC Candlestick",
    subtitle = "Four trading days",
    x = "Date",
    y = "Price"
  ) +
  theme_minimal()

p
```

### Candlestick with Moving Averages and Volume (ggplot2)

This example exercises the full accessible price + MA + volume pipeline:
a candlestick layer, two `geom_ma()` overlays (5- and 10-day SMAs), and
a separate volume bar panel composed via patchwork. MAIDR collapses the
two `geom_ma()` overlays into a single multi-series line layer, and the
candlestick + bar + line panels collapse to a single navigable subplot
in which each candle also carries its `volume` field.

```{r}
library(tidyquant)
library(patchwork)

set.seed(42)
n_days <- 20
dates  <- seq(as.Date("2024-01-02"), by = "day", length.out = n_days)
opens  <- 100 + cumsum(rnorm(n_days, 0, 1.5))
closes <- opens + rnorm(n_days, 0, 1.2)
highs  <- pmax(opens, closes) + abs(rnorm(n_days, 1, 0.5))
lows   <- pmin(opens, closes) - abs(rnorm(n_days, 1, 0.5))
vols   <- as.integer(runif(n_days, 1e5, 5e5))

ohlcv <- data.frame(
  date   = dates,
  open   = round(opens, 2),
  high   = round(highs, 2),
  low    = round(lows, 2),
  close  = round(closes, 2),
  volume = vols
)

p_price <- ggplot(
  ohlcv,
  aes(x = date, open = open, high = high, low = low, close = close)
) +
  geom_candlestick(
    colour_up = "darkgreen", colour_down = "red",
    fill_up = "darkgreen", fill_down = "red"
  ) +
  geom_ma(aes(y = close), ma_fun = SMA, n = 5,
          colour = "blue", linetype = "dashed", linewidth = 0.8) +
  geom_ma(aes(y = close), ma_fun = SMA, n = 10,
          colour = "orange", linetype = "dotted", linewidth = 0.8) +
  labs(title = "OHLC with 5- and 10-day SMA", x = NULL, y = "Price") +
  theme_minimal()

p_volume <- ggplot(ohlcv, aes(x = date, y = volume)) +
  geom_col(fill = "steelblue", alpha = 0.7) +
  labs(x = "Date", y = "Volume") +
  theme_minimal()

p_price / p_volume + plot_layout(heights = c(3, 1), axes = "collect_x")
```

### Base R Candlestick (quantmod)

The Base R path supports a plain OHLC candlestick via
`quantmod::chartSeries(x, type = "candlesticks")`. Each row of the
`xts`/`zoo` input is emitted as a navigable candle point with `value`
(ISO date), `open`, `high`, `low`, `close`, computed `trend`, and
`volatility`.

> **Limitations.** Technical-analysis overlays via the `TA` argument
> (e.g. `addVo()`, `addSMA()`, `addEMA()`) are **not supported** by the
> accessible HTML pipeline. The same applies to `chartSeries()`'s
> *default* `TA` whenever the input `xts` has a `Volume` column, since
> the default auto-adds `addVo()`. In all these cases maidr falls back
> to native (non-accessible) graphics with a one-time advisory pointing
> users to the ggplot2 + tidyquant + patchwork pipeline shown above.
> To opt in to accessible HTML for a Base R candlestick, supply OHLC
> data **without** a `Volume` column (default `TA` then becomes a
> no-op), or pass `TA = NULL` explicitly.

```{r}
#| eval: false
library(quantmod)

# OHLC-only xts (no Volume column) so the default TA is a no-op.
TST <- xts::xts(
  cbind(
    Open  = c(101.00, 102.00, 105.00, 103.50),
    High  = c(102.50, 105.50, 105.80, 104.50),
    Low   = c(100.50, 101.80, 103.00, 102.50),
    Close = c(102.00, 105.00, 103.50, 104.00)
  ),
  order.by = as.Date(c(
    "2024-01-12", "2024-01-13", "2024-01-14", "2024-01-15"
  ))
)
colnames(TST) <- c("TST.Open", "TST.High", "TST.Low", "TST.Close")

chartSeries(TST, type = "candlesticks", theme = "white", name = "TST")
```

---

## Heat Map

A heatmap uses color intensity to represent values in a two-dimensional
matrix. It is useful for spotting patterns, clusters, and outliers
across two categorical dimensions.

#### ggplot2

```{r}
heatmap_data <- expand.grid(
  Day = c("Mon", "Tue", "Wed", "Thu", "Fri"),
  Hour = c("9am", "10am", "11am", "12pm", "1pm", "2pm", "3pm", "4pm")
)
heatmap_data$Visitors <- c(
  20, 35, 50, 45, 30,
  25, 40, 60, 55, 35,
  30, 55, 75, 70, 45,
  40, 65, 90, 85, 60,
  35, 50, 70, 65, 40,
  30, 45, 60, 55, 38,
  25, 40, 55, 50, 32,
  15, 30, 40, 35, 22
)

p <- ggplot(heatmap_data, aes(x = Day, y = Hour, fill = Visitors)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "#f7fbff", high = "#08306b") +
  labs(title = "Website Visitors by Day and Hour") +
  theme_minimal()

p
```

#### Base R

```{r}
visitors <- matrix(
  c(
    20, 35, 50, 45, 30,
    25, 40, 60, 55, 35,
    30, 55, 75, 70, 45,
    40, 65, 90, 85, 60,
    35, 50, 70, 65, 40,
    30, 45, 60, 55, 38,
    25, 40, 55, 50, 32,
    15, 30, 40, 35, 22
  ),
  nrow = 5
)
image(visitors,
  col = hcl.colors(20, "Blues"),
  main = "Website Visitors by Day and Hour",
  xlab = "Day",
  ylab = "Hour",
  axes = FALSE
)
axis(1,
  at = seq(0, 1, length.out = 5),
  labels = c("Mon", "Tue", "Wed", "Thu", "Fri")
)
axis(2,
  at = seq(0, 1, length.out = 8),
  labels = c("9am", "10am", "11am", "12pm", "1pm", "2pm", "3pm", "4pm")
)
```

---

## KDE (Kernel Density Estimation) Plots

A density curve shows the estimated probability distribution of a
continuous variable. It provides a smooth alternative to histograms for
understanding the shape of data.

#### ggplot2

```{r}
density_values <- data.frame(value = c(
  rnorm(400, mean = 25, sd = 5),
  rnorm(600, mean = 40, sd = 8)
))

p <- ggplot(density_values, aes(x = value)) +
  geom_density(fill = "lightblue", alpha = 0.5, color = "steelblue") +
  labs(
    title = "Age Distribution of Survey Respondents",
    x = "Age",
    y = "Density"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
ages <- c(rnorm(400, mean = 25, sd = 5), rnorm(600, mean = 40, sd = 8))
d <- density(ages)
plot(d,
  col = "steelblue", lwd = 2,
  main = "Age Distribution of Survey Respondents",
  xlab = "Age",
  ylab = "Density"
)
polygon(d, col = rgb(0.678, 0.847, 0.902, 0.5), border = "steelblue")
```

---

## Regression Plots

A scatter plot with a fitted regression line shows the relationship
between two variables along with the best-fit model.

#### ggplot2

```{r}
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(
    title = "Weight vs MPG with Linear Fit",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
plot(mtcars$wt, mtcars$mpg,
  pch = 19, col = "steelblue",
  main = "Weight vs MPG with Linear Fit",
  xlab = "Weight (1000 lbs)", ylab = "Miles per Gallon"
)
abline(lm(mpg ~ wt, data = mtcars), col = "red", lwd = 2)
```

---

## Multi-Layered Plots

Multi-layered plots combine multiple visualization types in a single
chart. For example, a histogram overlaid with a density curve, or a bar
chart with a line overlay.

### Histogram with Density Overlay

#### ggplot2

```{r}
p <- ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(
    aes(y = after_stat(density)),
    bins = 15, fill = "lightblue", color = "white"
  ) +
  geom_density(color = "red", linewidth = 1.2) +
  labs(title = "MPG: Histogram with Density Curve") +
  theme_minimal()

p
```

#### Base R

```{r}
hist(mtcars$mpg,
  breaks = 15, freq = FALSE,
  col = "lightblue", border = "white",
  main = "MPG: Histogram with Density Curve",
  xlab = "Miles per Gallon",
  ylab = "Density"
)
lines(density(mtcars$mpg), col = "red", lwd = 2)
```

### Bar Chart with Line Overlay

#### ggplot2

```{r}
combo_data <- data.frame(
  month = factor(month.abb[1:6], levels = month.abb[1:6]),
  sales = c(100, 120, 90, 150, 130, 160),
  target = c(110, 110, 110, 140, 140, 140)
)

p <- ggplot(combo_data, aes(x = month)) +
  geom_bar(aes(y = sales), stat = "identity", fill = "steelblue", alpha = 0.7) +
  geom_line(aes(y = target, group = 1), color = "red", linewidth = 1.5) +
  labs(title = "Monthly Sales vs Target", y = "Value") +
  theme_minimal()

p
```

---

## Multi-Panel Plots (Multiple Subplots)

Multi-panel layouts arrange several independent plots in a grid. This is
useful for dashboards or comparing different views of the same dataset.

#### ggplot2 (patchwork)

```{r}
library(patchwork)

# Line plot with currency formatting
line_df <- data.frame(
  Month = 1:8,
  Revenue = c(2500, 4200, 3100, 5500, 4300, 6700, 5600, 7800)
)
pw_line <- ggplot(line_df, aes(Month, Revenue)) +
  geom_line(color = "steelblue", linewidth = 1) +
  labs(title = "Monthly Revenue", x = "Month", y = "Revenue") +
  theme_minimal()

# Bar plot with rates
bar_df1 <- data.frame(
  Category = c("A", "B", "C", "D", "E"),
  Rate = c(0.15, 0.22, 0.18, 0.28, 0.17)
)
pw_bar1 <- ggplot(bar_df1, aes(Category, Rate)) +
  geom_bar(stat = "identity", fill = "forestgreen", alpha = 0.7) +
  labs(title = "Conversion Rates", x = "Category", y = "Rate") +
  theme_minimal()

# Bar plot with large numbers
bar_df2 <- data.frame(
  Category = c("A", "B", "C", "D", "E"),
  Count = c(125000, 98000, 145000, 112000, 88000)
)
pw_bar2 <- ggplot(bar_df2, aes(Category, Count)) +
  geom_bar(stat = "identity", fill = "royalblue", alpha = 0.7) +
  labs(title = "User Counts", x = "Category", y = "Count") +
  theme_minimal()

# Line plot with exponential growth
line_df2 <- data.frame(
  x = 1:8,
  y = 10^(seq(3, 6.5, length.out = 8))
)
pw_line2 <- ggplot(line_df2, aes(x, y)) +
  geom_line(color = "tomato", linewidth = 1) +
  labs(title = "Exponential Growth", x = "Time", y = "Value") +
  theme_minimal()

combined <- (pw_line + pw_bar1 + pw_bar2 + pw_line2) +
  plot_layout(ncol = 2)
combined
```

#### Base R (par)

```{r}
par(mfrow = c(2, 2))

barplot(table(mtcars$cyl),
  col = "steelblue",
  main = "Cars by Cylinder Count",
  xlab = "Cylinders"
)

hist(mtcars$mpg,
  breaks = 12, col = "coral", border = "white",
  main = "MPG Distribution", xlab = "MPG"
)

plot(mtcars$wt, mtcars$mpg,
  pch = 19, col = "forestgreen",
  main = "MPG vs Weight",
  xlab = "Weight (1000 lbs)", ylab = "MPG"
)

boxplot(hp ~ gear,
  data = mtcars, col = "plum",
  main = "Horsepower by Gear Count",
  xlab = "Gears", ylab = "Horsepower"
)
invisible(par(mfrow = c(1, 1)))
```

---

## Facet Plots

Faceted plots split data into a grid of subplots by one or more grouping
variables. Each panel shows the same type of chart for a different
subset of the data. This is a ggplot2-only feature using `facet_wrap()`
or `facet_grid()`.

### facet_wrap

```{r}
facet_data <- data.frame(
  x = rep(c("A", "B", "C", "D"), 4),
  y = c(
    30, 25, 35, 20,
    45, 30, 25, 40,
    20, 35, 30, 45,
    35, 40, 20, 30
  ),
  panel = rep(c("Group 1", "Group 2", "Group 3", "Group 4"), each = 4)
)

p <- ggplot(facet_data, aes(x = x, y = y)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  facet_wrap(~panel, ncol = 2) +
  labs(title = "Sales by Category Across Groups", x = "Category", y = "Sales") +
  theme_minimal()

p
```

### facet_grid

```{r}
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(size = 2, color = "steelblue") +
  facet_grid(vs ~ am,
    labeller = labeller(
      vs = c("0" = "V-engine", "1" = "Straight"),
      am = c("0" = "Automatic", "1" = "Manual")
    )
  ) +
  labs(
    title = "MPG vs Weight by Engine Type and Transmission",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon"
  ) +
  theme_minimal()

p
```
