---
title: "Histogram, Density, Box and Violin Examples"
description: "Accessible histograms, kernel density (KDE) curves and box plots in ggplot2 and Base R, plus ggplot2 violin plots, with maidr: navigate bins, quartiles and density curves by keyboard with sonification."
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include = FALSE}
library(maidr)
library(ggplot2)
maidr_on()
set.seed(42)
```

Distribution plots describe the shape of one variable. A histogram is
walked bin by bin, a density curve point by point along the smoothed
curve, and a box plot as its five-number summary plus outliers, with
sonification tracing the rise and fall of each. Histograms, box plots and
density curves are stable plot types in both ggplot2 and Base R, and violin
plots are stable in ggplot2 (`geom_violin()`); the Base R reading of
`vioplot::vioplot()` is an experimental plot type (see "Supported plot
types" and "Experimental Plot Types" in the README), so this page shows
only the ggplot2 violin example. The [examples hub](examples.html) lists
every other plot family.

## Histogram

A histogram shows the frequency distribution of a continuous variable by
grouping values into bins. The height of each bar indicates how many
observations fall within that range.

### ggplot2

```{r}
exam_scores <- data.frame(score = c(
  rnorm(500, mean = 72, sd = 10),
  rnorm(500, mean = 85, sd = 8)
))

p <- ggplot(exam_scores, aes(x = score)) +
  geom_histogram(bins = 25, fill = "skyblue", color = "white") +
  labs(
    title = "Distribution of Exam Scores",
    x = "Score",
    y = "Frequency"
  ) +
  theme_minimal()

p
```

### Base R

```{r}
scores <- c(rnorm(500, mean = 72, sd = 10), rnorm(500, mean = 85, sd = 8))
hist(scores,
  breaks = 25,
  col = "skyblue",
  border = "white",
  main = "Distribution of Exam Scores",
  xlab = "Score",
  ylab = "Frequency"
)
```

---

## KDE (Kernel Density Estimation) Plots

A density curve shows the estimated probability distribution of a
continuous variable. It provides a smooth alternative to histograms for
understanding the shape of data.

### ggplot2

```{r}
density_values <- data.frame(value = c(
  rnorm(400, mean = 25, sd = 5),
  rnorm(600, mean = 40, sd = 8)
))

p <- ggplot(density_values, aes(x = value)) +
  geom_density(fill = "lightblue", alpha = 0.5, color = "steelblue") +
  labs(
    title = "Age Distribution of Survey Respondents",
    x = "Age",
    y = "Density"
  ) +
  theme_minimal()

p
```

### Base R

```{r}
ages <- c(rnorm(400, mean = 25, sd = 5), rnorm(600, mean = 40, sd = 8))
d <- density(ages)
plot(d,
  col = "steelblue", lwd = 2,
  main = "Age Distribution of Survey Respondents",
  xlab = "Age",
  ylab = "Density"
)
polygon(d, col = rgb(0.678, 0.847, 0.902, 0.5), border = "steelblue")
```

---

## Box Plot

A box plot summarizes a distribution using its five-number summary:
minimum, first quartile (Q1), median (Q2), third quartile (Q3), and
maximum. Outliers appear as individual points.

### ggplot2

```{r}
p <- ggplot(iris, aes(x = Species, y = Sepal.Length)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7) +
  labs(
    title = "Sepal Length by Iris Species",
    x = "Species",
    y = "Sepal Length (cm)"
  ) +
  theme_minimal()

p
```

### Base R

```{r}
boxplot(Sepal.Length ~ Species,
  data = iris,
  col = "lightblue",
  main = "Sepal Length by Iris Species",
  xlab = "Species",
  ylab = "Sepal Length (cm)"
)
```

---

## Violin Plot

A violin plot combines kernel density estimation (KDE) curves with
box-summary statistics, providing a richer view of the data distribution
than a box plot alone. MAIDR renders each violin as two navigable
layers: a **box** layer (min, Q1, median, Q3, max) and a **KDE** layer
(density curve).

> **Note:** Stable violin support is ggplot2 only. `vioplot::vioplot()` is
> read as an experimental Base R type; see the experimental table in the
> README.

```{r}
p <- ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_violin(fill = "lightblue", alpha = 0.7) +
  labs(
    title = "MPG Distribution by Cylinder Count",
    x = "Cylinders",
    y = "Miles per Gallon"
  ) +
  theme_minimal()

p
```
