---
title: "Experimental Base R Chart Examples"
description: "Accessible readings of experimental Base R charts with maidr: biplot(), stars() radar charts, mosaicplot(), cdplot(), filled.contour() and more, each mapped onto a navigable, sonified layer type."
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include = FALSE}
library(maidr)
library(ggplot2)
maidr_on()
set.seed(42)
```

Base R draws a long tail of statistical charts, from a biplot to a mosaic
plot, that maidr reads by mapping each onto a layer type it already
knows: a star plot becomes a radar with one series per row, a mosaic a
proportion grid, an association plot a heat map, so arrow-key navigation
and sonification carry over from those types. Every reading on this page
is an experimental plot type (see "Experimental Plot Types" in the
README): none has been through a user study and each may change without a
deprecation period. The [examples hub](examples.html) lists every other
plot family.

> **Note:** Everything below is Base R, so no ggplot2 counterpart is
> shown. Each section names the reading its chart gets, because the mapping
> is not always the one the chart's name suggests: an association plot is
> read as a heat map, a conditional density plot as a 100% stacked area.

## Biplot

`biplot()` puts observations and variable loadings on one pair of principal
component axes. maidr reads the observations as a **point** layer on `PC1`
and `PC2`, and each point carries its own row name, so a reader hears which
observation they are on rather than a bare coordinate pair.

```{r}
biplot(prcomp(USArrests, scale. = TRUE), main = "US arrests: PC1 vs PC2")
```

## Radar

`stars()` draws one star per row, with a ray per variable. maidr reads it as
a **radar**: one series per observation, each carrying a value per variable.

> **Note:** A star plot has no x and y axes to name, so the reading carries
> no axis labels — the variable names come through as the rays themselves.

```{r}
stars(mtcars[1:5, 1:4], main = "Five cars over four measures")
```

## Interaction Plot

`interaction.plot()` draws a line per level of the trace factor, so a
non-parallel pair is the interaction. maidr reads one **line** per level and
puts `trace.label` on the z axis, which is what names the two series apart.

```{r}
interaction.plot(ToothGrowth$dose, ToothGrowth$supp, ToothGrowth$len,
  xlab = "Dose (mg/day)", ylab = "Mean tooth length", trace.label = "Supplement"
)
```

## Box Plot from Summary Statistics

`bxp()` draws a box plot from statistics already computed, rather than from
raw data — which is exactly what `boxplot(plot = FALSE)` hands back. maidr
reads it as a **box**, the same reading `boxplot()` gets, so a pre-summarised
box is not a second-class one.

```{r}
bxp(boxplot(count ~ spray, data = InsectSprays, plot = FALSE),
  main = "Insect counts by spray"
)
```

## Strip Chart

`stripchart()` is the one-dimensional scatter you reach for when a box plot
would hide too few points. maidr reads **one point layer per group**, so the
groups are navigated as separate series rather than flattened together.

```{r}
stripchart(count ~ spray,
  data = InsectSprays, method = "jitter",
  xlab = "Count", ylab = "Spray"
)
```

## Dot Chart

`dotchart()` is Cleveland's alternative to a bar chart for labelled values.
maidr reads it as a **dot** layer.

```{r}
dotchart(VADeaths[, "Rural Male"],
  xlab = "Deaths per 1000", ylab = "Age group",
  main = "Virginia death rates, rural males"
)
```

## Lollipop

`plot(type = "h")` draws a vertical spike down to each value. maidr reads it
as a **lollipop**, the same type the correlogram uses.

```{r}
plot(1:8, c(2, 5, 3, 9, 4, 7, 6, 8),
  type = "h", xlab = "Index", ylab = "Value", main = "Spikes"
)
```

## Mosaic and Spine Plots

Both tile a contingency table so that area is proportion, and maidr reads
both as a **mosaic**: the y axis is `Proportion`, and the second factor is
named on z. `mosaicplot()` takes a two-way table.

```{r}
mosaicplot(HairEyeColor[, , "Male"], main = "Hair and eye colour, males")
```

`spineplot()` takes a factor response against one predictor, which makes it
the two-column case of the same reading.

```{r}
spineplot(factor(am) ~ wt, data = mtcars, xlab = "Weight", ylab = "Transmission")
```

## Conditional Density

`cdplot()` is the continuous-predictor counterpart to a spine plot: it draws
how the conditional distribution of a factor shifts along a numeric axis.
maidr reads it as a **100% stacked area**, since every vertical slice sums
to one.

```{r}
cdplot(factor(am) ~ mpg, data = mtcars, xlab = "Miles per gallon")
```

## Association Plot

`assocplot()` states one signed Pearson residual per cell of a contingency
table — how far that cell sits from independence. maidr reads it as a
**heat**: a named grid navigated row then column, with the residual on z, so
the sign and size of each departure are read out per cell.

```{r}
assocplot(HairEyeColor[, , "Male"], main = "Hair and eye colour: residuals")
```

## Filled Contour

`filled.contour()` shades the bands between contour lines. maidr reads the
**contour** levels themselves, so a reader walks the level curves rather
than the shading.

> **Note:** Passing `x` and `y` is worth doing here. Given only `z`,
> `filled.contour()` positions the grid on 0–1 and the reading has no axis
> labels at all; given real coordinates it announces them in their own
> units.

```{r}
# Every third row and column of the 87 x 61 grid: the level curves read the
# same, and the page carries about a ninth of the polygons.
volcano_coarse <- volcano[
  seq(1, nrow(volcano), by = 3), seq(1, ncol(volcano), by = 3)
]
filled.contour(
  x = 30 * seq_len(nrow(volcano_coarse)),
  y = 30 * seq_len(ncol(volcano_coarse)),
  z = volcano_coarse,
  xlab = "Easting (m)", ylab = "Northing (m)", main = "Maunga Whau"
)
```

## 100% Stacked Bar

A `barplot()` of a proportion table draws columns that each sum to one.
maidr reads it as a **100% stacked bar**, so a segment is announced as its
share rather than as a raw count.

```{r}
barplot(prop.table(table(mtcars$cyl, mtcars$gear), 2),
  xlab = "Gears", ylab = "Proportion", legend.text = TRUE
)
```
