---
title: "Scatter, Line, Step and Regression Examples"
description: "Accessible scatter plots, single and multi-line plots, step plots and regression fits in ggplot2 and Base R with maidr, each point or sample navigable by keyboard and audible through sonification."
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include = FALSE}
library(maidr)
library(ggplot2)
maidr_on()
set.seed(42)
```

Scatter, line and step plots are the x-y family: each point or sample is
one navigable term, Up and Down move between series in a multi-line
chart, and sonification plays the y value as pitch as you move along x. A
regression plot layers a fitted smooth over the points, so the fit is its
own navigable layer. Scatter, line, step and smooth layers are all stable
plot types (see "Supported plot types" in the README). The
[examples hub](examples.html) lists every other plot family.

## Scatter Plot

A scatter plot displays the relationship between two continuous
variables. Each point represents one observation, and color can encode a
third categorical variable.

### ggplot2

```{r}
car_data <- data.frame(
  weight = mtcars$wt,
  mpg = mtcars$mpg,
  cylinders = factor(mtcars$cyl)
)

p <- ggplot(car_data, aes(x = weight, y = mpg, color = cylinders)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    title = "Fuel Efficiency vs Vehicle Weight",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon",
    color = "Cylinders"
  ) +
  theme_minimal()

p
```

### Base R

```{r}
colors <- c("4" = "steelblue", "6" = "coral", "8" = "forestgreen")
plot(mtcars$wt, mtcars$mpg,
  pch = 19,
  col = colors[as.character(mtcars$cyl)],
  main = "Fuel Efficiency vs Vehicle Weight",
  xlab = "Weight (1000 lbs)",
  ylab = "Miles per Gallon"
)
legend("topright",
  legend = c("4 cyl", "6 cyl", "8 cyl"),
  col = colors, pch = 19
)
```

---

## Line Plots

### Single Line

A line plot connects ordered data points, making it ideal for showing
trends over time or sequential categories.

#### ggplot2

```{r}
temp_data <- data.frame(
  Month = factor(month.abb, levels = month.abb),
  Temperature = c(2, 4, 10, 15, 20, 25, 28, 27, 22, 15, 8, 3)
)

p <- ggplot(temp_data, aes(x = Month, y = Temperature, group = 1)) +
  geom_line(color = "tomato", linewidth = 1.2) +
  labs(
    title = "Average Monthly Temperature",
    y = "Temperature (C)"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
months <- 1:12
temps <- c(2, 4, 10, 15, 20, 25, 28, 27, 22, 15, 8, 3)
plot(months, temps,
  type = "l", col = "tomato", lwd = 2,
  main = "Average Monthly Temperature",
  xlab = "Month", ylab = "Temperature (C)",
  xaxt = "n"
)
axis(1, at = 1:12, labels = month.abb)
```

### Multiple Lines

A multi-line plot overlays several series on the same axes, enabling
direct comparison of trends across groups.

#### ggplot2

```{r}
multi_line <- data.frame(
  Year = rep(2015:2024, 3),
  Users = c(
    10, 15, 22, 35, 50, 72, 95, 120, 150, 180,
    8, 12, 18, 25, 38, 55, 70, 88, 110, 135,
    5, 8, 14, 20, 30, 42, 58, 75, 95, 118
  ),
  Platform = rep(c("Mobile", "Desktop", "Tablet"), each = 10)
)

p <- ggplot(multi_line, aes(x = Year, y = Users, color = Platform)) +
  geom_line(linewidth = 1.2) +
  labs(
    title = "Platform Users Over Time",
    y = "Users (millions)"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
years <- 2015:2024
users <- cbind(
  Mobile  = c(10, 15, 22, 35, 50, 72, 95, 120, 150, 180),
  Desktop = c(8, 12, 18, 25, 38, 55, 70, 88, 110, 135),
  Tablet  = c(5, 8, 14, 20, 30, 42, 58, 75, 95, 118)
)

matplot(years, users,
  type = "l", lwd = 2,
  col = c("steelblue", "coral", "forestgreen"),
  lty = 1,
  main = "Platform Users Over Time",
  xlab = "Year", ylab = "Users (millions)"
)
legend("topleft",
  legend = colnames(users),
  col = c("steelblue", "coral", "forestgreen"),
  lwd = 2
)
```

---

## Step Plot

A step plot suits a value that is **piecewise constant**: it is held across an
interval and then jumps, rather than drifting between samples the way a line
implies. The canonical case is a **hypnogram** — the sleep stage a sleep study
scores for each epoch of the night.

MAIDR reports the layer's step convention (`hv`, `vh`, or `mid`) so the
description says where the value jumps, and offers a **Transitions** rotor mode
that moves between the moments the level changes rather than sample by sample.

When the y aesthetic is an ordinal factor, the level *name* is announced
("REM", "N2", "Awake") while the numeric level still drives sonification and
braille — so the shape of the night is audible and the stage is speakable.

### ggplot2

```{r}
hypnogram <- data.frame(
  hour = seq(0, 7.5, by = 0.5),
  stage = factor(
    c(
      "Awake", "N1", "N2", "N3", "N3", "N2", "REM", "N2",
      "N3", "N3", "N2", "REM", "N2", "N1", "REM", "Awake"
    ),
    levels = c("N3", "N2", "N1", "REM", "Awake")
  )
)

p <- ggplot(hypnogram, aes(x = hour, y = stage, group = 1)) +
  geom_step(direction = "hv", color = "steelblue", linewidth = 1) +
  scale_x_continuous(breaks = 0:8) +
  labs(
    title = "Overnight Hypnogram",
    x = "Hours after lights out",
    y = "Sleep stage"
  ) +
  theme_minimal()

p
```

`geom_step(direction = )` accepts `"hv"` (the default — hold, then jump at the
next x), `"vh"` (jump at the current x, then hold), and `"mid"` (jump midway
between x values). All three are passed through to MAIDR unchanged.

### Base R

`plot(type = "s")` draws the horizontal segment first, matching `"hv"`;
`plot(type = "S")` draws the vertical segment first, matching `"vh"`.

```{r}
hours <- seq(0, 7.5, by = 0.5)
stage_names <- c("N3", "N2", "N1", "REM", "Awake")
stage <- c(
  5, 3, 2, 1, 1, 2, 4, 2,
  1, 1, 2, 4, 2, 3, 4, 5
)

plot(hours, stage,
  type = "s", col = "steelblue", lwd = 2,
  main = "Overnight Hypnogram",
  xlab = "Hours after lights out", ylab = "Sleep stage",
  yaxt = "n"
)
axis(2, at = seq_along(stage_names), labels = stage_names, las = 1)
```

---

## Regression Plots

A scatter plot with a fitted regression line shows the relationship
between two variables along with the best-fit model.

### ggplot2

```{r}
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(color = "steelblue", size = 3) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(
    title = "Weight vs MPG with Linear Fit",
    x = "Weight (1000 lbs)",
    y = "Miles per Gallon"
  ) +
  theme_minimal()

p
```

### Base R

```{r}
plot(mtcars$wt, mtcars$mpg,
  pch = 19, col = "steelblue",
  main = "Weight vs MPG with Linear Fit",
  xlab = "Weight (1000 lbs)", ylab = "Miles per Gallon"
)
abline(lm(mpg ~ wt, data = mtcars), col = "red", lwd = 2)
```
