---
title: "Bar, Pie and Word Cloud Examples"
description: "Accessible bar charts (simple, dodged and stacked), pie charts and word clouds in ggplot2 and Base R with maidr: keyboard navigation, screen reader text and sonification for each category."
execute:
  echo: true
  warning: false
  message: false
---

```{r setup, include = FALSE}
library(maidr)
library(ggplot2)
maidr_on()
set.seed(42)
```

Bar and pie charts are the categorical family: one navigable term per bar,
wedge or word, moved through with the Left and Right arrow keys, with the
value of each announced and mapped to pitch when sonification is on. Bar
and pie charts are stable plot types in both ggplot2 and Base R (see
"Supported plot types" in the README); the word cloud is an experimental
Base R type whose reading may change without a deprecation period. The
[examples hub](examples.html) lists every other plot family.

## Bar Plots

### Simple Bar Chart

A simple bar chart compares values across categories. Each bar
represents one category with its height proportional to its value.

#### ggplot2

```{r}
bar_data <- data.frame(
  Product = c("Laptop", "Tablet", "Phone", "Monitor"),
  Sales = c(150, 230, 180, 290)
)

p <- ggplot(bar_data, aes(x = Product, y = Sales)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Product Sales by Category",
    x = "Product",
    y = "Sales (units)"
  ) +
  theme_minimal()

p
```

#### Base R

```{r}
products <- c("Laptop", "Tablet", "Phone", "Monitor")
sales <- c(150, 230, 180, 290)

barplot(sales,
  names.arg = products,
  col = "steelblue",
  main = "Product Sales by Category",
  xlab = "Product",
  ylab = "Sales (units)"
)
```

### Dodged / Grouped Bar Chart

A dodged bar chart places bars for each sub-group side by side, making
it easy to compare values within and across categories.

#### ggplot2

```{r}
dodged_data <- data.frame(
  Region = rep(c("North", "South", "East"), each = 2),
  Quarter = rep(c("Q1", "Q2"), 3),
  Revenue = c(120, 150, 200, 180, 160, 210)
)

p <- ggplot(dodged_data, aes(x = Region, y = Revenue, fill = Quarter)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(title = "Quarterly Revenue by Region") +
  scale_fill_manual(values = c("steelblue", "coral")) +
  theme_minimal()

p
```

#### Base R

```{r}
revenue_matrix <- matrix(c(120, 150, 200, 180, 160, 210), nrow = 2)
rownames(revenue_matrix) <- c("Q1", "Q2")
barplot(revenue_matrix,
  beside = TRUE,
  names.arg = c("North", "South", "East"),
  col = c("steelblue", "coral"),
  legend.text = rownames(revenue_matrix),
  main = "Quarterly Revenue by Region",
  xlab = "Region",
  ylab = "Revenue"
)
```

### Stacked Bar Chart

A stacked bar chart layers sub-groups on top of each other within each
category, showing both individual contributions and the total.

#### ggplot2

```{r}
stacked_data <- data.frame(
  Year = rep(c("2022", "2023", "2024"), each = 3),
  Source = rep(c("Solar", "Wind", "Hydro"), 3),
  Output = c(40, 30, 50, 55, 45, 48, 70, 60, 52)
)

p <- ggplot(stacked_data, aes(x = Year, y = Output, fill = Source)) +
  geom_bar(stat = "identity", position = position_stack()) +
  labs(
    title = "Renewable Energy Output by Source",
    y = "Output (GWh)"
  ) +
  scale_fill_manual(values = c("#2ecc71", "#3498db", "#9b59b6")) +
  theme_minimal()

p
```

#### Base R

```{r}
energy_matrix <- matrix(
  c(40, 30, 50, 55, 45, 48, 70, 60, 52),
  nrow = 3
)
rownames(energy_matrix) <- c("Solar", "Wind", "Hydro")
barplot(energy_matrix,
  beside = FALSE,
  names.arg = c("2022", "2023", "2024"),
  col = c("#2ecc71", "#3498db", "#9b59b6"),
  legend.text = rownames(energy_matrix),
  main = "Renewable Energy Output by Source",
  xlab = "Year",
  ylab = "Output (GWh)"
)
```

---

## Pie Chart

A pie chart splits a whole into wedges, one per category, each wedge's
angle proportional to its share. maidr announces every slice by name and
value, and derives its percentage of the total for you.

### ggplot2

ggplot2 has no pie geom: a pie is a single stacked column bent around by
`coord_polar("y")`, so `x` is the literal `""` and the categories go on
`fill`. `coord_radial(theta = "y")` works the same way. Note that
`coord_polar("x")` draws a coxcomb instead, which maidr still describes as a
bar chart.

```{r}
market_share <- data.frame(
  Browser = c("Chrome", "Safari", "Edge", "Firefox"),
  Share = c(64, 19, 5, 3)
)

p <- ggplot(market_share, aes(x = "", y = Share, fill = Browser)) +
  geom_col(width = 1) +
  coord_polar("y") +
  labs(
    title = "Browser Market Share",
    fill = "Browser",
    y = "Share (%)"
  ) +
  theme_void()

p
```

### Base R

```{r}
shares <- c(Chrome = 64, Safari = 19, Edge = 5, Firefox = 3)
pie(shares,
  col = c("#3498db", "#2ecc71", "#9b59b6", "#e67e22"),
  main = "Browser Market Share",
  xlab = "Browser",
  ylab = "Share (%)"
)
```

---

## Word Cloud

A word cloud is the extreme case of a chart that carries real data while
being readable only by eye: each term's weight is drawn as **glyph size**
and written down nowhere on the page. Structurally it is a categorical
label and a magnitude, so maidr reads it as a term and its number.

> **Experimental.** `word_cloud` is one of the experimental layer types.
> It has not been through a user study, and it may change without a
> deprecation period. See the experimental table in the README.

> **Note:** Word clouds are a Base R feature. ggplot2 has no word cloud
> geom, and the packages that add one draw through their own devices
> rather than through a layer maidr can read.

> **Requires the {wordcloud} package.** Install with
> `install.packages("wordcloud")`.

### Base R

The counts you pass survive into the reading, so the weight axis honestly
says **Occurrences**. (The Python binding cannot: `wordcloud.WordCloud`
divides every frequency by the largest and keeps only the ratio, so
py-maidr announces a *relative* frequency. Same chart, two different
honest readings.)

```{r}
mentions <- c(
  accessibility = 412, sonification = 300, braille = 250,
  screenreader = 190, keyboard = 155, contrast = 120
)

maidr::wordcloud(
  words = names(mentions),
  freq = mentions,
  min.freq = 1,
  random.order = FALSE,
  colors = c(
    "#3498db", "#2ecc71", "#9b59b6",
    "#e67e22", "#e74c3c", "#16a085"
  )
)
```

> **Attach order matters.** `library(wordcloud)` after `library(maidr)`
> puts `package:wordcloud` ahead of `package:maidr` on the search path, so
> a bare `wordcloud()` call reaches the wordcloud package directly and
> maidr never records it — the chart draws, but `show()` and
> `save_html()` then report that no Base R plot was found. Measured on a
> three-term cloud:
>
> | How it is called | Recorded? |
> |---|---|
> | `library(maidr)` then `library(wordcloud)`, bare `wordcloud()` | **no** |
> | `library(wordcloud)` then `library(maidr)`, bare `wordcloud()` | yes |
> | `maidr::wordcloud()` | yes |
>
> Either attach `wordcloud` **before** `maidr`, or call
> `maidr::wordcloud()` explicitly as above, which works in either order.

> **No highlighting.** `wordcloud()` draws each term with a bare `text()`
> call at a rotation chosen by `rot.per`, and nothing names those — the
> exported SVG carries no `id` attributes at all. So a word cloud is read
> without a visual highlight, rather than pairing the terms with whatever
> else happened to resolve.

Two of `wordcloud()`'s arguments decide **which** terms are drawn, and the
reading replicates both so it never announces a term the chart left out:
`min.freq` (default 3) drops anything rarer, and `max.words` keeps only
the heaviest. `wordcloud()` also lowers `min.freq` to 0 when it exceeds
every frequency, which is what stops a cloud of rare terms coming out
empty; that rule is copied rather than approximated.
